\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-25 coqimpl.tex
Minor Fix. Please Check coqimpl.
Fix meta-theory example.
edits & new comments

Note that, in our example, we include two
functors without arguments---this is due to a peculiarity of Coq's
"Include": non-instantiated parameters will be automatically
instantiated by the appropriate fields in the surrounding defining
module.
For example, "b_7" requires a module of type "a_6". Inside module "A",
when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat}
inside the current surrounding "A". Then when we
\mintinline{Coq}{Include b_7} without specifying the module parameter,
Coq will try to find fields in the surrounding module (i.e., module
"A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by
the \mintinline{Coq}{a : nat} that is included earlier.
The resulting compiled module is just like mundane Coq's module and 
thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.

\begin{figure}[!htb]
  \begin{minipage}[t]{0.45\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Family I.
Field a : nat := 1.
Field b : a = a := eq_refl.
Fail Field c : a = 1 := eq_refl. (*Fail 1*)

Final Field a' : nat := 1.
Field c : a = 1 := eq_refl.
Field add : nat -> nat -> nat := (+).
Fail Field comm_add : ...

@@

EndFamily.
\end{minted}
  \end{minipage}
\begin{minipage}[t]{0.45\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Family I2 extends I.
Override Field a : nat := 1. (* Override *)
Inherit b.                   (* Inherit  *)

@@

Override Field add := (-).

Inherits Until MonotonicF.
Override Family MonotonicF.
  Final Field f : nat := (+ 1).
  Field p : monotonic f := ...
SealFamily. 
EndFamily.
\end{minted}
  \end{minipage}
  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
\end{figure}

% explaining inheritance
Besides standalone family, we also need to implement family inheritance.
To simplify the implementation, we can consider all families as
inheritance and the standalone families are extending the empty family.
Thus we only need to deal with inheritance during interactive theorem proving.

There are only three kinds of inheritance we are dealing with---extend, override, and inherit. We use first-order data to encode inheritance data inside the implementation of the plugin. The inheritance data will include the information of (1) the ``type'' of the parent family, (2) a list of operation indicating how each field of the parent will be dealt with (either inherited or overrided), (3) the operations indicating newly extended fields, and (4) the ``type'' of the children family.\YZ{what is 'inheritance data'?}\EDJreply{It is just inheritance judgement, but before meta-theory is introduced, I can only say some "first-order data that represent the information of inheritance". I emphasize first order to distinguish from functor in ML community.}\YZreply{I think this is the place where you can use an example to demonstrate these three kinds of inheritance.} The surface syntax for the three kinds of inheritance are demonstrated in \cref*{fig:plugin-example3}. The implementation of overriding, inheritance, and extension is achieved by simply manipulating the internal representation of a family correspondingly---we can swap the module in the list for overriding, retain the module in the list for inheritance and add new module into the list for extension. The compilation from a family (internal representation) to a module still acts the same.

Since it is possible to consider inheritance as solely a first-order data rather
than a higher-order functor, it is actually plausible to
\text{mix-in} two inheritance data -- by carefully ``mix'' the internal list.

\textbf{Reasoning Requires Non-Overridable.}\YZ{At this point in this section, overriding has not been discussed yet, but it is already discussing non-overridables. Should probably discuss the extensibility first.}
This compilation strategy
has already allowed fields to be type-dependent on former fields thus fulfilling
\ref{langdesign-req1}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
the concrete definition of a given field (especially when we are proving
properties about it).

% For example, we may
% define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
% and intend to prove its commutativity in a following field $"comm"$.
% But to prove so, we need to know more than just a type interface of
% "add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
% future, once "add" is overridden into a subtraction function then
% "comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
% A more concrete example is in \cref{fig:plugin-example1}, consider the
% case we have "Field b : a = 1 := .." instead. Then our
% compilation and Coq's type checking will both fail: in the context of
% "b" (module type "a_6") we only know "a" is of type "nat" but nothing
% more (and this is necessary so that "a" can be overridden in the
% future).
Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the example, we can see that the first attempt of proving "a = 1" fails. This should fail, actually---because we want to retain the possibility to override "a" in the future, but the mere existence of the field "c" will stop that, since "c" witnesses "a" fix to "1". From the other perspective, the proof of "c" will require a concrete definition of "a" instead of just the exposure of its type. \textbf{We can see a conflict between \textit{overridable} and \textit{(definition-relevant) reasoning}}.  We enable the exposure of the concrete definition by providing "Final" keyword upon "a'".

This example might sounds trivial and non-practical. Let's look at the example field "add". Now that we only expose the type of "add", we cannot prove its commutativity in "comm_add" (because we don't have access to its concrete definition) but we can override "add" with subtraction function in "Family I2". This makes sense because we shouldn't have (inherited) commutativity property for subtraction anyway. This is contrary to how "a'" is declared "Final", the second "c" can prove that it's equal to 1 but "a'" is not overridable in the "Family I2".
\YZ{Not clear if you want to be able to override a or want to be able to have 'Field b : a = 1'.}\EDJreply{Rephrased. Please Check.}

@@

To resolve this, we provide family sealing mechanism to simulate sigma type---we allow the
users to specify the compiled module type of a given family,\YZ{I don't understand "specify the family/compiled module type"}\EDJreply{Sorry. I rephase it into "compiled module type of a given family".}\YZreply{Still cannot see what is going on and how it is related to the sentence that follows. An example is due.}\EDJreply{Done. Please check.}
and thus we can have a
 \textit{non-overridable} (thus locally transparent) field located in
an \textit{overridable} family. Check the "MonotonicF" in \cref{fig:plugin-example3}, we simply seal
\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
together inside an overridable family that does not allow "f"
to be overridden in any of its derived families.
Because "f" is non-overridable inside the family, its definition is
transparent and can be used to prove "p".
However, because the enclosing family of "f" is overridable, to swap the
implementation of "f", we just override the whole ``sealed'' family, like we do in "I2.MonotonicF".

In our example, the seal family "I.MonotonicF" is actually simulating interfaces as well\YZ{I was actually thinking sealed families simulate sigma types, no?}\EDJreply{You are absolutely right. I should add a "can" here. I can use sealed family to simulate interface is by creating a sigma type, but making the implemntation of sigma type trivial (using place holder/using axiom as I specified below). Then once I want to instantiate with real implementation, I override the whole family -- as long as it has the same type as the sealed family, everything still works.}\YZreply{Can you explain this with an example? I recall the AbsInt example uses it?}\EDJreply{Done. Please check.}. We have \\ ``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}'', and we have assigned it to both "f" and "p", and later instantiate in "I2" with concrete implementation by overriding this family "I.MonotonicF".  \YZ{I don't understand what it means to "assign each field with an axiom".}\EDJreply{It really means ``place holder'' in general programming language (Python's NonImplement()), defer to future. Because Coq's Axiom can be used to instantiate anything.}
% Basically, {a : int -> int; property : a is monotonic}, in this case
%   we may want to override "a" using different computation, but we still wnat
%   this refinement/property

>>>>

Note that, in our example, we include two
functors without arguments---this is due to a peculiarity of Coq's
"Include": uninstantiated parameters will be automatically
instantiated by the appropriate fields in the surrounding defining
module.
For example, "b_7" requires a module of type "a_6". Inside module "A",
when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat}
inside the current surrounding "A". Then when we
\mintinline{Coq}{Include b_7} without specifying the module parameter,
Coq will try to find fields in the surrounding module (i.e., module
"A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by
the \mintinline{Coq}{a : nat} that is included earlier.
The resulting compiled module is just like mundane Coq's module and 
thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.

\begin{figure}[!htb]
  \begin{minipage}[t]{0.45\linewidth}
\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
Family I.
Field a : nat := 1.
Field b : a = a := eq_refl.
Fail Field c : a = 1 := eq_refl. (*Fail 1*)

Final Field a' : nat := 1.
Field c : a@\YZ{should this be a'?}@ = 1 := eq_refl.
Field add : nat -> nat -> nat := (+).
Fail Field comm_add : ...

@@

EndFamily.
\end{minted}
  \end{minipage}
\begin{minipage}[t]{0.45\linewidth}
\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
Family I2 extends I.
Override Field a : nat := 1. (* Override *)
Inherit b.                   (* Inherit  *)

@@

Override Field add := (-).@\YZ{'add' is not a good choice of name---it doesn't make sense to override add with subtraction.}@

Inherits Until MonotonicF.
Override Family MonotonicF.
  Final Field f : nat@\YZ{nat -> nat ?}@ := (+ 1).
  Field p : monotonic f := ...
SealFamily. 
EndFamily.
\end{minted}
  \end{minipage}
  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
\end{figure}

% explaining inheritance
Besides standalone family, we also need to implement family inheritance.
To simplify the implementation, we can consider all families as
inheritance and the standalone families are extending the empty family.
Thus, we need to deal with only inheritance during interactive theorem proving.

There are three kinds of inheritance---extend, override, and inherit. We
use first-order data to encode inheritance data inside the
implementation of the plugin. The inheritance data will include the
information of (1) the ``type'' of the parent family, (2) a list of
operation indicating how each field of the parent will be dealt with
(either inherited or overriden), (3) the operations indicating newly
extended fields, and (4) the ``type'' of the children family.
The surface syntax for the three kinds of inheritance
are demonstrated in the exemplar \cref*{fig:plugin-example3}. The
implementation of overriding, inheritance, and extension is achieved by
simply manipulating the internal representation of a family
correspondingly---we can swap the module in the list for overriding,
retain the module in the list for inheritance, and add new module into
the list for extension. The compilation from a family (internal
representation) to a module still acts the same.

Since it is possible to consider inheritance as solely a first-order data rather
than a higher-order functor, it is possible to
\text{mix-in} two inheritance data---by carefully ``mix'' the internal list.\YZ{How about making mixin composition one of your Req's? Also does any example in the paper illustrate how to do this?}

\textbf{Reasoning Requires Non-Overridable.}
This compilation strategy
has already allowed fields to be type-dependent on former fields thus fulfilling
\ref{langdesign-req1}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
the concrete definition of a given field (especially when we are proving
properties about it).

% For example, we may
% define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
% and intend to prove its commutativity in a following field $"comm"$.
% But to prove so, we need to know more than just a type interface of
% "add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
% future, once "add" is overridden into a subtraction function then
% "comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
% A more concrete example is in \cref{fig:plugin-example1}, consider the
% case we have "Field b : a = 1 := .." instead. Then our
% compilation and Coq's type checking will both fail: in the context of
% "b" (module type "a_6") we only know "a" is of type "nat" but nothing
% more (and this is necessary so that "a" can be overridden in the
% future).
Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the
example. We can see that the first attempt of proving "a = 1" fails.
This should fail, actually---because we want to retain the possibility
to override "a" in the future, but the mere existence of the field "c"
will stop that, since "c" witnesses "a" fix to "1". From the other
perspective, the proof of "c" will require a concrete definition of "a"
instead of just the exposure of its type. \textbf{We can see a conflict
between \textit{overridable} and \textit{(definition-relevant)
reasoning}}.  We enable the exposure of the concrete definition by
providing "Final" keyword upon "a'".

This example might sound trivial and contrived. Let's look at the
example field "add". Now that we only expose the type of "add", we
cannot prove its commutativity in "comm_add" (because we don't have
access to its concrete definition) but we can override "add" with
subtraction function in "Family I2". This makes sense because we
shouldn't have (inherited) commutativity property for subtraction
anyway. This is contrary to how "a'" is declared "Final", the second "c"
can prove that it's equal to 1 but "a'" is not overridable in the
"Family I2".

@@

To resolve this, we provide family sealing mechanism to simulate sigma type---we allow the
users to specify the compiled module type of a given family, and thus we
can have a \textit{non-overridable} (thus locally transparent) field
located in an \textit{overridable} family.
See, for example, the "MonotonicF" in \cref{fig:plugin-example3}.
We simply seal
\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
together inside an overridable family that does not allow "f"
to be overridden in any of its derived families.
Because "f" is non-overridable inside the family, its definition is
transparent and can be used to prove "p".
However, because the enclosing family of "f" is overridable, to swap the
implementation of "f", we just override the whole ``sealed'' family,
like we do in "I2.MonotonicF". And the compiled module type of
"MonotonicF" will be fixed throughout the extension in different
families---it is always these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}

In our example, the seal family "I.MonotonicF" is actually simulating
interfaces as well.  We have
\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
and we have assigned it to both "f" and "p", and later instantiate in
"I2" with concrete implementation by overriding this family
"I.MonotonicF".
% Basically, {a : int -> int; property : a is monotonic}, in this case
%   we may want to override "a" using different computation, but we still wnat
%   this refinement/property
